---
title: "Fitting Step-Selection Functions with `amt`"
author: "Johannes Signer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting a Step-Selection Function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
set.seed(20161113)
```


## About

This vignette briefly introduces how one can fit a Step-Selection Function (SSF) with the `amt` package. We will be using the example data of one red deer from northern Germany and one covariate: a forest cover map.

## Getting the data ready

First we load the required libraries and the relocation data (called `deer`)

```{r}
library(lubridate)
library(raster)
library(amt)
data("deer")
deer
```

In order to continue, we need a regular sampling rate. To check the current sampling rate, we use `summarize_sampling_rate`:

```{r}
summarize_sampling_rate(deer)
```

The median sampling rate is 6h, which is what we aimed for.


Next, we have to get the environmental covariates. A forest layer is included in the package. Note, that this a regular `RasterLayer`. 

```{r}
data("sh_forest")
sh_forest
```

## Prepare Data for SSF

Before fitting a step selection, the data well need to prepared. First, we change from a point representation to a step representation, using the function `steps_by_burst`, which in contrast to the `steps` function accounts for bursts.

```{r}
ssf1 <- deer %>% steps_by_burst()
```

Next, we generate random steps with the function `random_steps`. This function fits by default a Gamma distribution to the step lengths and a von Mises distribution to the turn angles, and then pairs each observed step with `n` random steps.

```{r}
ssf1 <- ssf1 %>% random_steps(n = 15)
```

As a last step, we have to extract the covariates at the end point of each step. We can do this with `extract_covariates`.

```{r}
ssf1 <- ssf1 %>% extract_covariates(sh_forest) 
```


Since the forest layers is coded as `1 = forest` and `2 != forest`, we create a factor with appropriate levels. We also calculate the log of the step length and the cosine of the turn angle, which we may use later for a integrated step selection function.


```{r}
ssf1 <- ssf1 %>% 
  mutate(forest = factor(sh.forest, levels = 1:2, labels = c("forest", "non-forest")), 
         cos_ta = cos(ta_), 
         log_sl = log(sl_)) 
```

## Fitting SSF 

Now all pieces are there to fit a SSF. We will use `fit_clogit`, which is a wrapper around `survival::clogit`.

```{r}
m0 <- ssf1 %>% fit_clogit(case_ ~ forest + strata(step_id_))
m1 <- ssf1 %>% fit_clogit(case_ ~ forest + forest:cos_ta + forest:log_sl + log_sl * cos_ta + strata(step_id_))
m2 <- ssf1 %>% fit_clogit(case_ ~ forest + forest:cos_ta + forest:log_sl + log_sl + cos_ta + strata(step_id_))
summary(m0)
summary(m1)
summary(m2)
AIC(m0$model)
AIC(m1$model)
AIC(m2$model)
```

### Interpretation of coefficients

To be done.

<!---
The coefficients are:

```{r}
coef(m2)
```


- `forestnon-forest`: There is positive selection for non-forested areas, although it is not significant (at a $\alpha = 0.05$ level). Exponentiating the selection coefficient for forest, results with a relative selection strength of `r round(exp(coef(m2)[1]))`.




- The coefficients for `log_sl` and `cos_ta` are both signicant and indicating that the intial parameters need to be adjusted.
- The interaction between 
- 
- 

--->

## A note on piping

All steps described above, could easily be wrapped into one piped workflow:

```{r}
m1 <- deer %>% 
  steps_by_burst() %>% random_steps(n = 15) %>% 
  extract_covariates(sh_forest) %>% 
  mutate(forest = factor(sh.forest, levels = 1:2, labels = c("forest", "non-forest")), 
         cos_ta = cos(ta_), 
         log_sl = log(sl_)) %>% 
  fit_clogit(case_ ~ forest + forest:cos_ta + forest:sl_ + sl_ * cos_ta + strata(step_id_))
```

```{r}
summary(m1)
```


## Session

```{r}
devtools::session_info()
```

